\textbf{\Large Hybrid Client/Server Rendering with Automatic Proxy Model Generation}\\[0.2ex]

### 2017-10-18 tex/full_2017/paper2.tex
Attempt at accommodating reviewers and their comments/suggestions.

\newcommand{\ie}{{\em i.e.}}
\newcommand{\eg}{{\em e.g.}}
\newcommand{\etal}{{\em et~al.}}

@@

\author{\IEEEauthorblockN{~\\[-0.4ex]\large Jens Olav Nygaard\\[0.3ex]\normalsize}
\IEEEauthorblockA{SINTEF Digital\\
Applied Mathematics\\
Norway\\
Email: {\tt Jens.O.Nygaard@sintef.no}}
\and
\IEEEauthorblockN{~\\[-0.4ex]\large Jon Mikkelsen Hjelmervik\\[0.3ex]\normalsize}
\IEEEauthorblockA{SINTEF Digital\\
Applied Mathematics\\
Norway\\
Email: {\tt Jon.M.Hjelmervik@sintef.no}}}

@@

\twocolumn[{%
\renewcommand\twocolumn[1][]{#1}%
\maketitle
\begin{center}
    \centering
		\begin{tabular}[b]{ccc}
    \includegraphics[width=0.32\linewidth, trim={1cm 1cm 1cm 1cm}, clip]{frviewb1.png} &
    \includegraphics[width=0.32\linewidth, trim={0 1cm 0 1.5cm}, clip]{frviewb2.png} &
    \includegraphics[width=0.32\linewidth, trim={0 1.5cm 0 2cm}, clip]{frviewb3.png} \\
		a) & b) & c)
		\end{tabular}
		\captionof{figure}{\label{fig:FRView} Oil reservoir viewer showing one
  server-rendered and two client-rendered images of automatically generated and
  slightly rotated proxy models.  For~(b), parameters are chosen to produce the
  best possible image, and in~(c) we want to highlight artifacts and
  implementational details. See the text for further discussion.}
%    \protect \label{fig:wfedesign} }
%    \captionof{figure}{Test caption}
\end{center}%
}]

@@

\begin{abstract}
%\boldmath
We describe an algorithm and implementation of automatic proxy model generation
for a client/server remote rendering setup. A {\em proxy model} is a
light-weight version of a main 3D model that is cheaper to render and
transfer. We automatically derive this from the main model. The client is
typically a tablet, for which we only assume availability of WebGL and
Javascript. The server makes use of OpenGL and a web server. When the rate of
received server-rendered images deteriorate, the client renders the proxy model,
which is computed from depth buffers bundled with rendered images from the
server. This algorithm requires little modification of the application itself.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

@@

Hand in hand with increasingly powerful rendering engines comes ever increasing
requirements on computational accuracy, power efficiency, data sizes, scaling
properties, etc. This is also reinforced by popular cloud-based approaches and
wireless usage patterns. An effect of this is that interactivity still is a
difficult issue. We consider a client/server model for 3D rendering, addressing
latency, bandwidth and scaling problems in a novel way.

We introduce a {\em proxy model}, defined as a temporary model to be shown and
manipulated locally on a client while waiting for the appropriate image from a
connected server. Producing such proxy models can be difficult for many kinds of
3D data, like in our main case in which we have an oil reservoir viewer
(YYYY,~\cite{cloudviz}) that renders large {\em corner point grids}, together
with faults, oil wells, and more. See Figure~\ref{fig:FRView} above for an
example of both a full server-side rendering, and automatically generated proxy
models rendered in Google Chrome.
%
Our solution is to pass depth information from the server along with ordinary
rendered images. From this, a rudimentary 3D model is built, and with the RGB
image as a texture, this model can be manipulated and rendered on the client
while waiting for the next update from the server. If the client does not change
the position or orientation of the model too much, this proxy model rendering
integrates seamlessly with the slower stream of server-rendered frames.
%
Even if bandwidth and latency is not a problem, it may be desirable to let a
server of limited capacity serve many simultaneous users, hence limiting the
effective server time available for each one. Suitable scaling may still be
achieved using our solution. This is currently being commercialized as a part of
the result of  a recently finished EU-project called CloudFlow.

%-------------------------------------------------------------------------
\subsection{Previous work}
\label{sec:prevWork}

Our approach has some similarities to {\em image based rendering} (IBR)
techniques, with the difference that an important IBR problem would be the
reconstruction of a depth map from images, while we have access to the full
depth map from a rendering pass.  Another way to use depth maps similar to what
we do, is for ``immersive streaming'', see \eg,~\cite{ibr}, where focus is on
depth map compression, an issue that we also consider. Another work
focused on similar streaming and geometry compression,
is~\cite{teler}. Also,~\cite{220764} contains some of the same ideas as our
work, with respect to image-based rendering acceleration.
%
\textbf{A very early work aiming at the same kind of ``inter-frame rendering'', is
\cite{Mark:1997:PW:253284.253292}. Here, several frames are warped, and
subsequently combined, in order to avoid large unpainted areas caused by
occlusion. This corresponds to our use of several proxy models, each from a pair
of (rgb, depth)-images, the main difference being that they work on separate
pixels, while we use larger, textured ``splats''. Our method avoids their
slightly complicated meshing, discontinuity detection and final image
composition stage.}
%
Common to many IBR-algorithms is also that of stitching together 3D or 2.5D
point clouds. We could do this for our 2.5D maps fetched from the server, but it
is unclear if the benefit would outweigh the cost.
%
Our approach differs from many similar ones, in that we use existing depth maps 
to distill and render temporary geometry, rather than retrieveing the depth maps from images.
%
We observe that these 2.5D height maps are exactly what ``3D cameras''
(time-of-flight and other range image sensors) produce, but most authors
considering these are typically building more complex geometries before
visualization, see~\cite{IMM2009-05801}.
%
\textbf{Another approach is that of~\cite{altenhofen16rixels}. They use websockets
where we use the http protocol, a more significant difference is that we get
away with transferring a lot less data due to our adaptive compression ratio
depending on continuous bandwidth and latency measurements.
%
}
%
\textbf{In~\cite{CGF:CGF1871}, Pajak \etal considers remote rendering and
  streaming of frames rendered from a dynamical 3D model, we are quite agnostic
  to the source of imagery. Their setup requires more powerful clients than ours
  (OpenGL vs. WebGL), but they will also have higher fidelity.  }
%
\textbf{A special case is provided for in the rendering of stereo images. In
  this case, the 3D disparity is limited, while the rendering cost is doubled,
  since two views per frame is needed. In~\cite{DidykERMS2010}, this is
  exploited to make a solution tailored to such stereo synthesis; performance
  approaches that of rendering non-stereo, with a minimization of depth
  disparity artifacts. Occlusion and disocclusion holes are avoided by warping quads rather than
  pixel and filling in with previous images.}
%
% Er denne egentlig relevant?
%\textbf{Nalbach \etal,~\cite{Nalbach:2014:DSS:2556700.2556708}, ...
%}

%-------------------------------------------------------------------------
\section{The auto proxy algorithm}

@@

\begin{figure}[htb]
  \centering
  \subfigure[Low frame rate]{
    \input{fig3a.tex}
  }
  \subfigure[Initial proxy model]{
    \input{fig3b.tex}
  }
  \subfigure[Client-generated frame]{
    \input{fig3c.tex}
  }
  \subfigure[With texturing]{
    \input{fig3d.tex}
  }
  \subfigure[Screen-space-sized splats]{
    \input{fig3e.tex}
  }
  \subfigure[With intra-splat depths]{
    \input{fig3f.tex}
  }
  \caption{\label{fig:2DheightmapRotated}
Bird's view of 3D model (solid lines) and proxy model renderings.
%           The depth map and image sent from the server enables the client to
%           recover a 3D model approximating the one rendered on the server. This
%           client-side model is shown here with a small additional rotation, again
%           seen from above.
}
\end{figure}

@@

%-------------------------------------------------------------------------
\subsection{The server side}

@@

When the client receives an RGB and depth image, together with view
transformations, it builds a proxy model from this. This model can then be
transformed and rendered directly, or it can be combined with other proxy models
the client already has in store, see Section~\ref{sec:proxyModelReplacement}.

As indicated by Figure~\ref{fig:2DheightmapRotated}~(b), the received height map
does not allow us more than concluding where a set of points belong on the 3D
model, \ie, we have little topologic information. Since the information from the
depth map can only contain the foremost point along any ray from the observer,
it is said to be in 2.5D, as opposed to 3D.  The simplest thing the client can
do, is just to transform and render the set of 3D points with color sampled from
the server-rendered image, as is illustrated in
Figure~\ref{fig:2DheightmapRotated}~(b).

Rendering a 3D point for each depth fragment available may tax a thin client,
and it may be necessary to use a smaller number of primitives and instead render
each with a larger number of pixels on the client side, a {\em splat}.  A set of
such splats for a given depth image we call a {\em proxy model}. By rendering
splats, we get fewer ``false connections'' than if we connect 3D points and
reconstruct topology, but we risk getting more ``holes'' in our models.

We can render each splat as a fixed geometry, \eg, a disk or rectangle, with the
correspondig color from the image, see Figure~\ref{fig:2DheightmapRotated}~(c),
where we have introduced a transformation local to the client.  We will briefly
describe some improvements to this. Note that other possibilities include
building and maintaining a 3D occupancy mesh, computing a distance field from
which iso-surfaces can be extracted, etc.

@@

Each splat produces many client pixels, necessitating an ``intra-splat''
fragment texturing for which we need a local texture coordinate transform.
A first approximation is for the client to assume that the corresponding part of
the server's model is planar in a region around the given point. If this is the
case, a local 2D texture transformation will provide a good approximation to the
intra-splat texturing to be performed on the client. Let $P_c$ and $P_s$ be
projection matrices on the client and server, respectively, and $M_c$ and $M_s$
corresponding view matrices. For the splat centered in $\vv_{i, j} = (x_j, y_i,
z_{i, j})$, to be centered on the client's canvas at screen coordinate $\pv_{i,
j} = P_c M_c M_s^{-1} P_s^{-1}\vv_{i, j} = U\vv_{i, j}$, the texture coordinate
transformation to be used is,
\[
  T =
  \begin{pmatrix}
    \frac{1}{n_x} & 0 \\
    0 & \frac{1}{n_y}
  \end{pmatrix}
  \Big( \sv_x \, \, \, \sv_y \Big)^{-1}
  \begin{pmatrix} 
    \frac{w}{n_x} & 0 \\
     0 & \frac{h}{n_y}
  \end{pmatrix}
   =
  A
  \Big( \sv_x \, \, \, \sv_y \Big)^{-1}
  B,
\]
where $A$ maps the ``client's splat region'' (in $[0, 1]^2$) to the
corresponding texture area, $(\sv_{x} \, \, \, \sv_{y})^{-1}$ maps the
``client's screen space splat area'' to $[0, 1]^2$, and $B$ provides a scaling
factor to fill the {\em viewport} of size $w \times h$ with $n_x \times n_y$
splats laid out in a grid, when $M_c=M_s$.
%
% After mult by proj matrix: clip coo, range of (x, y, z) = [-w, w]^3
% After persp division: ndc [-1, 1]^3
% After viewport transf: window coordinates
% (Range [vp[0] vp[1]] x [[vp[2] vp[3]] x [0, 1] ?)
%
We obtain $\sv_{x}$ and $\sv_{y}$ by evaluating proxy model positions followed by 
perspective division and transformation into window coordinates,
\[
  \pv =
  U { \small \begin{pmatrix} x \\ y \\ d \\ 1 \end{pmatrix} },
  \pv_{x+\Delta x} =
  U { \small \begin{pmatrix} x+\Delta x \\ y \\ d_{\Delta x} \\ 1 \end{pmatrix} }
  \text{and}\, \, \, 
  \pv_{y+\Delta y} =
  U { \small \begin{pmatrix} x \\ y+\Delta y \\ d_{\Delta y} \\ 1 \end{pmatrix} },
\]
where $d = 2D(x, y) - 1$, $d_{\Delta x} =
2D(x+\Delta x, y) - 1$, and $d_{\Delta y} = 2D(x,
y+\Delta y) - 1$ are depths sampled from the received depth buffer
$D$ and transformed to $[-1, 1]$. With a
$w$-component set to one, we have in effect done a perspective division, so that
we are in clip space and multiplication with $U$ is appropriate.
%
Note that $\Delta x$ and $\Delta y$ are not unique, we use
\[
  \Delta x = n_x / \text{width}(D) \text{\,\,\,\,and\,\,\,\,}
  \Delta y = n_y / \text{height}(D),
%  \label{eq:deltaChoice}
\]
but something larger may also be used. These are used for looking up depth image
samples, and the calculation of $\sv_x$ and $\sv_y$ is really a gradient
approximation, so we should not make them too large either.

This leaves us $\pv$, $\pv_{x+\Delta x}$ and $\pv_{y+\Delta y}$ in clip
coordinates, and from this we get
corresponding window coordinates $\sv$, $\sv_{x+\Delta x}$ and $\sv_{y+\Delta
y}$, from which we finally get splat spanning vectors
\[
  \sv_{\delta} =
  \sv_{x+\delta} - \sv =
    \frac{L}{2\delta} \left(
        \frac{\pv_{x+\delta}.xy}{\pv_{x+\delta}.w} -
        \frac{\pv.xy}{\pv.w}
    \right),
\]
with $\delta=\Delta x$ or $\delta=\Delta y$, $L$ is viewport size, either $w$ or
$h$, and we have used the ``shader notation'' for vector components.  The client
renders a large \texttt{glPoint} for each splat, with texture coordinates $(s,
t)^\prime$, and each fragment then looks up the server-rendered image at
position
\[
  \begin{pmatrix}
    s \\ t
  \end{pmatrix} +
  T 
  \begin{pmatrix}
    u \\ v
  \end{pmatrix}
  =
  \begin{pmatrix}
    s \\ t
  \end{pmatrix} +
  T 
  \begin{pmatrix}
    \texttt{glPointCoord}-\frac{1}{2}
  \end{pmatrix}
\]
where $(u, v)$ are ``intra-splat texture coordinates''.  When the assumption
that the geometry is locally planar does not hold, \eg, if the splats are very
large, or they originate from a curved or non-smooth part, this may look rather
odd, see Figure~\ref{fig:LargeSplatsOnCorners}.

\begin{figure}[htb]
  \centering
  \subfigure[Server-rendered]{
    \includegraphics[width=.3\linewidth]{splat1.jpg}
  }
  \subfigure[$P_c M_c = P_s M_s$]{
    \includegraphics[width=.3\linewidth]{splat2.jpg}
  }
  \subfigure[$P_c M_c \neq P_s M_s$]{
    \includegraphics[width=.3\linewidth]{splat3.jpg}
  }
  \caption{\label{fig:LargeSplatsOnCorners}
           Splats for which depths are sampled on a planar region look nice on
  this region, but possibly contorted elsewhere. Notice the corner and edges
  here, where we have chosen parameters to display this effect.}
\end{figure}

@@

It is important that the process of generating and sending proxy model data from
the server itself does not slow down transmission.  There are mainly three
sources of delay for server-rendered images to the client; high latency, low
bandwidth, and slow server-rendering.  In the first case, it seems prudent to
have a good proxy model on the client, which can be used for longer time and for
a wider range of client-side transformations.  In the two other cases, it is
important for the proxy model generation/transmission to be cheap/fast, both in order
to get the proxy model to the client and keep from delaying the server-image
more than necessary. These demands are not always compatible.

We have adopted an adaptive specification of proxy model data from the server
involving a more light-weight image (lossy JPG-compression with adaptive quality
control) while interaction is ongoing.  Also, reduced resolution of the depth
buffer sent from the server is used. Another possibility is to let the client
dynamically set the number of splats, number of proxy models, etc.

%-------------------------------------------------------------------------
\section{Results and discussion}

>>>>

%\newcommand{\ie}{{\em i.e.}}
%\newcommand{\eg}{{\em e.g.}}
%\newcommand{\etal}{{\em et~al.}}
\newcommand{\ie}{{i.e.}}
\newcommand{\eg}{{e.g.}}
\newcommand{\etal}{{et~al.}}

@@

%\author{\IEEEauthorblockN{~\\[-0.4ex]\large Jens Olav Nygaard\\[0.3ex]\normalsize}
%\IEEEauthorblockA{SINTEF Digital\\
%Applied Mathematics\\
%Norway\\
%Email: {\tt Jens.O.Nygaard@sintef.no}}
%\and
%\IEEEauthorblockN{~\\[-0.4ex]\large Jon Mikkelsen Hjelmervik\\[0.3ex]\normalsize}
%\IEEEauthorblockA{SINTEF Digital\\
%Applied Mathematics\\
%Norway\\
%Email: {\tt Jon.M.Hjelmervik@sintef.no}}}

\author{
\IEEEauthorblockN{~\\[-0.4ex]\large Jens Olav Nygaard\\ and Jon Mikkelsen Hjelmervik\\[0.3ex]\normalsize}
\IEEEauthorblockA{SINTEF Digital\\
Applied Mathematics\\
Norway\\
Email: {\tt Jens.O.Nygaard@sintef.no},\\ and {\tt Jon.M.Hjelmervik@sintef.no}}
}

%\author{
%\IEEEauthorblockN{~\\[-0.4ex]\large Michael Shell\\[0.3ex]\normalsize}
%\IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
%Georgia Institute of Technology\\
%Atlanta, Georgia 30332--0250\\
%Email: {\tt ...}}
%\and
%\IEEEauthorblockN{~\\[-0.4ex]\large Homer Simpson\\[0.3ex]\normalsize}
%\IEEEauthorblockA{Twentieth Century Fox\\
%Springfield, USA\\
%Email: {\tt homer@thesimpsons.com}}
%\and
%\IEEEauthorblockN{~\\[-0.4ex]\large James Kirk\\ and Montgomery Scott\\[0.3ex]\normalsize}
%\IEEEauthorblockA{Starfleet Academy\\
%San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212\\
%Fax: (888) 555--1212}
%}

@@

\twocolumn[{%
\renewcommand\twocolumn[1][]{#1}%
\maketitle
%\begin{center}
%    \centering
%		\begin{tabular}[b]{ccc}
%    \includegraphics[width=0.32\linewidth, trim={1cm 1cm 1cm 1cm}, clip]{frviewb1.png} &
%    \includegraphics[width=0.32\linewidth, trim={0 1cm 0 1.5cm}, clip]{frviewb2.png} &
%    \includegraphics[width=0.32\linewidth, trim={0 1.5cm 0 2cm}, clip]{frviewb3.png} \\
%		a) & b) & c)
%		\end{tabular}
%		\captionof{figure}{\label{fig:FRView} Oil reservoir viewer showing one
%  server-rendered and two client-rendered images of automatically generated and
%  slightly rotated proxy models.  For~(b), parameters are chosen to produce the
%  best possible image, and in~(c) we want to highlight artifacts and
%  implementational details. See the text for further discussion.}
%%    \protect \label{fig:wfedesign} }
%%    \captionof{figure}{Test caption}
%\end{center}%
}]

@@

\begin{abstract}
%\boldmath
%We describe an algorithm and implementation of automatic proxy model generation
%for a client/server remote rendering setup. A {\em proxy model} is a
%light-weight version of a main 3D model that is cheaper to render and
%transfer. We automatically derive this from the main model. The client is
%typically a tablet, for which we only assume availability of WebGL and
%Javascript. The server makes use of OpenGL and a web server. When the rate of
%received server-rendered images deteriorate, the client renders the proxy model,
%which is computed from depth buffers bundled with rendered images from the
%server. This algorithm requires little modification of the application itself.
%
%Revisit the Abstract to introduce the topic, as for a regular reader; currently the abstract starts too abruptly into the matter.  I suggest:
%a) What is the problem in a given domain?
%b) Why is the problem interesting? (a few sentences, shortly)
%c) What other did/failed on, and what you are dealing with
%d) What is your solution? (one sentence, very short)
%e) End the abstract with one sentence on what is the main output of the paper
%
%The algorithm is very well described and the computation-flow formulas for generating the proxy model are accurately described.
%
% New version after comments from reviewers
%
A common problem in remote rendering setups, is that of temporarily insufficient
bandwidth and latency. For a proper experience of immersiveness, at least some
rendering should be presented to the user, and it should appear to be responsive
to user input, even in the presence of connection glitches. The all too familiar
spinning hourglass symbol, or equivalent, will degrade such an experience. With
the advent of VR and AR, solutions to this become important even though
connectedness in general is improving. We are dealing with a remote rendering of
3D geometry being pushed to a lesser client, and what we in essence do is to
replace a spinning hourglass symbol with an automatically client-generated
approximation of the 3D geometry rendered on the client, responding to
client-recorded user input. We call this a {\em proxy model}.  Our main result
is a system for automatically producing such proxy models on the client, from
received images and depth buffers only, for showing on the client when remotely
rendered frames do not arrive sufficiently fast.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

@@

\begin{figure*}
\begin{center}
    \centering
		\begin{tabular}[b]{ccc}
    \includegraphics[width=0.32\linewidth, trim={1cm 1cm 1cm 1cm}, clip]{frviewb1.png} &
    \includegraphics[width=0.32\linewidth, trim={0 1cm 0 1.5cm}, clip]{frviewb2.png} &
    \includegraphics[width=0.32\linewidth, trim={0 1.5cm 0 2cm}, clip]{frviewb3.png} \\
		a) & b) & c)
		\end{tabular}
		\captionof{figure}{\label{fig:FRView} Oil reservoir viewer showing one
  server-rendered and two client-rendered images of automatically generated and
  slightly rotated proxy models.  For~(b), parameters are chosen to produce the
  best possible image, and in~(c) we want to highlight artifacts and
  implementational details. See the text for further discussion.}
\end{center}%
\end{figure*}

Hand in hand with increasingly powerful rendering engines comes ever increasing
requirements on computational accuracy, power efficiency, data sizes, scaling
properties, etc. This is also reinforced by popular cloud-based approaches and
wireless usage patterns. An effect of this is that interactivity still is a
difficult issue. We consider a client/server model for 3D rendering, addressing
latency, bandwidth and scaling problems in a novel way.
\\
We introduce a {\em proxy model}, defined as a temporary model to be shown and
manipulated locally on a client while waiting for the appropriate image from a
connected server. Producing such proxy models can be difficult for many kinds of
3D data, like in our main case in which we have an oil reservoir
viewer~\cite{cloudviz} that renders large {\em corner point grids}, together
with faults, oil wells, and more; see Figure~\ref{fig:FRView} for an
example of both a full server-side rendering, and automatically generated proxy
models rendered in Google Chrome.
%
Our solution is to pass depth information from the server along with ordinary
rendered images. From this, a rudimentary 3D model is built, and with the RGB
image as a texture, this model can be manipulated and rendered on the client
while waiting for the next update from the server. If the client does not change
the position or orientation of the model too much, this proxy model rendering
integrates seamlessly with the slower stream of server-rendered frames.
%
Even if bandwidth and latency is not a problem, it may be desirable to let a
server of limited capacity serve many simultaneous users, hence limiting the
effective server time available for each one. Suitable scaling may still be
achieved using our solution. This is currently being commercialized as a part of
the result of  a recently finished EU-project called CloudFlow~\cite{cloudflow}.
\\
In Section~\ref{sec:prevWork}, we review some previous and related work, before
we consider our contribution in Section~\ref{sec:autoproxy}. This is further
split into three subsections, in Subsection~\ref{subsec:server} we consider the
server side part of the system, in Subsection~\ref{sec:client}, the client side
and in Subsection~\ref{sec:autoTuning} we consider automatic parameter tuning
during use of the system. We discuss results in Section~\ref{sec:results},
before we finally sum up in Subsection~\ref{sec:conc}.

%-------------------------------------------------------------------------
\section{Previous work}
\label{sec:prevWork}

Our approach has some similarities to {\em Image Based Rendering} (IBR)
techniques~\cite{journalsritaOliveira02}, with the difference that an
important IBR problem would be the reconstruction of a depth map from images,
while we have access to the full depth map from a rendering pass.  Another way
to use depth maps similar to what we do, is for ``immersive streaming'', see,
\eg,~\cite{ibr}, where focus is on depth map compression, an issue that we also
consider. Another work focused on similar streaming and geometry compression,
is Teler~\cite{teler}. Also,~\cite{220764} contains some of the same ideas as our
work, with respect to image-based rendering acceleration.
%
A very early work aiming at the same kind of ``inter-frame rendering'', is
Mark \etal~\cite{Mark:1997:PW:253284.253292}. Here, several frames are warped,
and subsequently combined, in order to avoid large unpainted areas caused by
occlusion. This corresponds to our use of several proxy models, each from a pair
of (rgb, depth)-images, the main difference being that they work on separate
pixels, while we use larger, textured ``splats''. Our method avoids their
slightly complicated meshing, discontinuity detection and final image
composition stage.
%
Common to many IBR-algorithms is also that of stitching together 3D or 2.5D
point clouds. We could do this for our 2.5D maps fetched from the server, but it
is unclear if the benefit would outweigh the cost.
%
Our approach differs from many similar ones, in that we use existing depth maps 
to distill and render temporary geometry, rather than retrieveing the depth maps from images.
%
We observe that these 2.5D height maps are exactly what ``3D cameras''
(time-of-flight and other range image sensors) produce, but most authors
considering these are typically building more complex geometries before
visualization, see~\cite{IMM2009-05801}.
%
Another approach is that of~\cite{altenhofen16rixels}. They use websockets
where we use the http protocol, a more significant difference is that we get
away with transferring a lot less data due to our adaptive compression ratio
depending on continuous bandwidth and latency measurements.
%
In~\cite{CGF:CGF1871}, Pajak \etal~considers remote rendering and
streaming of frames rendered from a dynamical 3D model, we are quite agnostic
to the source of imagery. Their setup requires more powerful clients than ours
(OpenGL vs. WebGL), but they will also have higher fidelity.
%
A special case is provided for in the rendering of stereo images. In this case,
the 3D disparity is limited, while the rendering cost is doubled, since two
views per frame is needed. In~\cite{DidykERMS2010}, this is exploited to make a
solution tailored to such stereo synthesis; performance approaches that of
rendering non-stereo, with a minimization of depth disparity
artifacts. Occlusion and disocclusion holes are avoided by warping quads rather
than pixel and filling in with previous images.
%
% Er denne egentlig relevant?
%\textbf{Nalbach \etal,~\cite{Nalbach:2014:DSS:2556700.2556708}, ...
%}

%-------------------------------------------------------------------------
\section{The auto proxy algorithm}
\label{sec:autoproxy}

@@

\begin{figure}[htb]
  \centering
  \subfigure[Low frame rate]{
    \input{fig3a.tex}
  }
  \subfigure[Initial proxy model]{
    \input{fig3b.tex}
  }
  \subfigure[Client-generated frame]{
    \input{fig3c.tex}
  }
  \subfigure[With texturing]{
    \input{fig3d.tex}
  }
  \subfigure[Screen-space-sized splats]{
    \input{fig3e.tex}
  }
  \subfigure[With intra-splat depths]{
    \input{fig3f.tex}
  }
  \caption{\label{fig:2DheightmapRotated}
Bird's view of 3D model (solid lines) and proxy\\model renderings.
}
\end{figure}

%           The depth map and image sent from the server enables the client to
%           recover a 3D model approximating the one rendered on the server. This
%           client-side model is shown here with a small additional rotation, again
%           seen from above.

@@

%-------------------------------------------------------------------------
\subsection{The server side}
\label{subsec:server}

@@

When the client receives an RGB and depth image, together with view
transformations, it builds a proxy model from this. This model can then be
transformed and rendered directly, or it can be combined with other proxy models
the client already has in store, see Section~\ref{sec:proxyModelReplacement}.
\\
As indicated by Figure~\ref{fig:2DheightmapRotated}~(b), the received height map
does not allow us more than concluding where a set of points belong on the 3D
model, \ie, we have little topologic information. Since the information from the
depth map can only contain the foremost point along any ray from the observer,
it is said to be in 2.5D, as opposed to 3D.  The simplest thing the client can
do, is just to transform and render the set of 3D points with color sampled from
the server-rendered image, as is illustrated in
Figure~\ref{fig:2DheightmapRotated}~(b).
\\
Rendering a 3D point for each depth fragment available may tax a thin client,
and it may be necessary to use a smaller number of primitives and instead render
each with a larger number of pixels on the client side, a {\em splat}.  A set of
such splats for a given depth image we call a {\em proxy model}. By rendering
splats, we get fewer ``false connections'' than if we connect 3D points and
reconstruct topology, but we risk getting more ``holes'' in our models.
\\
We can render each splat as a fixed geometry, \eg, a disk or rectangle, with the
correspondig color from the image, see Figure~\ref{fig:2DheightmapRotated}~(c),
where we have introduced a transformation local to the client.  We will briefly
describe some improvements to this. Note that other possibilities include
building and maintaining a 3D occupancy mesh, computing a distance field from
which iso-surfaces can be extracted, etc.

@@

Each splat produces many client pixels, necessitating an ``intra-splat''
fragment texturing for which we need a local texture coordinate transform.
A first approximation is for the client to assume that the corresponding part of
the server's model is planar in a region around the given point. If this is the
case, a local 2D texture transformation will provide a good approximation to the
intra-splat texturing to be performed on the client. Let $P_c$ and $P_s$ be
projection matrices on the client and server, respectively, and $M_c$ and $M_s$
corresponding view matrices. For the splat centered in $\vv_{i, j} = (x_j, y_i,
z_{i, j})$, to be centered on the client's canvas at screen coordinate $\pv_{i,
j} = P_c M_c M_s^{-1} P_s^{-1}\vv_{i, j} = U\vv_{i, j}$, the texture coordinate
transformation to be used is,
\[
  T =
  \begin{pmatrix}
    \frac{1}{n_x} & 0 \\
    0 & \frac{1}{n_y}
  \end{pmatrix}
  \Big( \sv_x \, \, \, \sv_y \Big)^{-1}
  \begin{pmatrix} 
    \frac{w}{n_x} & 0 \\
     0 & \frac{h}{n_y}
  \end{pmatrix}
   =
  A
  \Big( \sv_x \, \, \, \sv_y \Big)^{-1}
  B,
\]
where $A$ maps the ``client's splat region'' (in $[0, 1]^2$) to the
corresponding texture area, $(\sv_{x} \, \, \, \sv_{y})^{-1}$ maps the
``client's screen space splat area'' to $[0, 1]^2$, and $B$ provides a scaling
factor to fill the {\em viewport} of size $w \times h$ with $n_x \times n_y$
splats laid out in a grid, when $M_c=M_s$.
%
% After mult by proj matrix: clip coo, range of (x, y, z) = [-w, w]^3
% After persp division: ndc [-1, 1]^3
% After viewport transf: window coordinates
% (Range [vp[0] vp[1]] x [[vp[2] vp[3]] x [0, 1] ?)
%
We obtain $\sv_{x}$ and $\sv_{y}$ by evaluating proxy model positions followed by 
perspective division and transformation into window coordinates,
\[
  \pv =
  U { \small \begin{pmatrix} x \\ y \\ d \\ 1 \end{pmatrix} },
  \pv_{x+\Delta x} =
  U { \small \begin{pmatrix} x+\Delta x \\ y \\ d_{\Delta x} \\ 1 \end{pmatrix} }
  \text{and}\, \, \, 
  \pv_{y+\Delta y} =
  U { \small \begin{pmatrix} x \\ y+\Delta y \\ d_{\Delta y} \\ 1 \end{pmatrix} },
\]
where $d = 2D(x, y) - 1$, $d_{\Delta x} =
2D(x+\Delta x, y) - 1$, and $d_{\Delta y} = 2D(x,
y+\Delta y) - 1$ are depths sampled from the received depth buffer
$D$ and transformed to $[-1, 1]$. With a
$w$-component set to one, we have in effect done a perspective division, so that
we are in clip space and multiplication with $U$ is appropriate.
%
Note that $\Delta x$ and $\Delta y$ are not unique, we use
\[
  \Delta x = n_x / \text{width}(D) \text{\,\,\,\,and\,\,\,\,}
  \Delta y = n_y / \text{height}(D),
%  \label{eq:deltaChoice}
\]
but something larger may also be used. These are used for looking up depth image
samples, and the calculation of $\sv_x$ and $\sv_y$ is really a gradient
approximation, so we should not make them too large either.
\\
This leaves us $\pv$, $\pv_{x+\Delta x}$ and $\pv_{y+\Delta y}$ in clip
coordinates, and from this we get
corresponding window coordinates $\sv$, $\sv_{x+\Delta x}$ and $\sv_{y+\Delta
y}$, from which we finally get splat spanning vectors
\[
  \sv_{\delta} =
  \sv_{x+\delta} - \sv =
    \frac{L}{2\delta} \left(
        \frac{\pv_{x+\delta}.xy}{\pv_{x+\delta}.w} -
        \frac{\pv.xy}{\pv.w}
    \right),
\]
with $\delta=\Delta x$ or $\delta=\Delta y$, $L$ is viewport size, either $w$ or
$h$, and we have used the ``shader notation'' for vector components.  The client
renders a large \texttt{glPoint} for each splat, with texture coordinates $(s,
t)^\prime$, and each fragment then looks up the server-rendered image at
position
\[
  \begin{pmatrix}
    s \\ t
  \end{pmatrix} +
  T 
  \begin{pmatrix}
    u \\ v
  \end{pmatrix}
  =
  \begin{pmatrix}
    s \\ t
  \end{pmatrix} +
  T 
  \begin{pmatrix}
    \texttt{glPointCoord}-\frac{1}{2}
  \end{pmatrix}
\]
where $(u, v)$ are ``intra-splat texture coordinates''.  When the assumption
that the geometry is locally planar does not hold, \eg, if the splats are very
large, or they originate from a curved or non-smooth part, this may look
slightly odd, see Figure~\ref{fig:LargeSplatsOnCorners}. The figure shows splats
for which depths are sampled on different planar regions at an angle to one
another, causing contortions when large splats cover more than one such planar
region.

\begin{figure}[htb]
  \centering
  \subfigure[Server-rendered]{
    \includegraphics[width=.3\linewidth]{splat1.jpg}
  }
  \subfigure[$P_c M_c = P_s M_s$]{
    \includegraphics[width=.3\linewidth]{splat2.jpg}
  }
  \subfigure[$P_c M_c \neq P_s M_s$]{
    \includegraphics[width=.3\linewidth]{splat3.jpg}
  }
  \caption{\label{fig:LargeSplatsOnCorners}
% Reviewer said ``too much text''...
%  Splats for which depths are sampled on a planar region look nice on
%  this region, but possibly contorted elsewhere. Notice the corner and edges
%  here, where we have chosen parameters to display this effect.}
  Notice the corner and edges, parameters are chosen to display effects of
  different planes meeting at edges.}
\end{figure}

@@

It is important that the process of generating and sending proxy model data from
the server itself does not slow down transmission.  There are mainly three
sources of delay for server-rendered images to the client; high latency, low
bandwidth, and slow server-rendering.  In the first case, it seems prudent to
have a good proxy model on the client, which can be used for longer time and for
a wider range of client-side transformations.  In the two other cases, it is
important for the proxy model generation/transmission to be cheap/fast, both in order
to get the proxy model to the client and keep from delaying the server-image
more than necessary. These demands are not always compatible.
\\
We have adopted an adaptive specification of proxy model data from the server
involving a more light-weight image (lossy JPG-compression with adaptive quality
control) while interaction is ongoing.  Also, reduced resolution of the depth
buffer sent from the server is used. Another possibility is to let the client
dynamically set the number of splats, number of proxy models, etc.

%-------------------------------------------------------------------------
\section{Results and discussion}
\label{sec:results}
